Notes

### 2022-09-19 chapters/appendix-code.tex
Small changes after revision

The Python source code developed for this work is available at the GitHub repository \url{https://github.com/antcc/rk-bfr}. The code is adequately documented and is structured in several directories as follows:

\begin{itemize}
  \item In the \texttt{rkbfr} folder we find the files responsible for the implementation of our models, separated according to the functionality they provide.
  \item The \texttt{reference\_methods} folder contains our implementation of the functional comparison algorithms that were not available through a standard Python library.
  \item At the root folder we have files for executing our experiments, which accept many user-specified parameters (e.g. number of iterations, type of data set, \ldots), along with some utility files. In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while script the \texttt{results\_all.py} allows the execution of our models without a cross-validation loop.
\end{itemize}

When possible, the code was implemented in a generic way that would allow for easy extensions or derivations. It was also developed with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy} and \textit{scipy} libraries. Moreover, we followed closely the style of the \textit{scikit-learn} and \textit{scikit-fda} libraries, so our methods are fully compatible and could be integrated (after some minor tweaking) with both of them. The code for the experiments was executed in a machine with 4 cores, and a random seed with value 2022 was set for reproducibility. We provide a script file \texttt{launch.sh} that illustrates a typical execution. Lastly, there are also Jupyter notebooks that demonstrate the use of our methods in a more visual way. Inside these notebooks there is a step-by-step guide on how one might execute our algorithms, accompanied by many graphical representations, and offering the possibility of changing multiple parameters to experiment with the code. In addition, there is also a notebook that can be used to generate all the tables and figures of this document pertaining to the experimental results.

>>>>

The Python source code developed for this work is available at the GitHub repository \url{https://github.com/antcc/rk-bfr}, under a GPLv3 license. The code is adequately documented and is structured in several directories as follows:

\begin{itemize}
  \item In the \texttt{rkbfr} folder we find the files responsible for the implementation of our models, separated according to the functionality they provide.
  \item The \texttt{reference\_methods} folder contains our implementation of the functional comparison algorithms that were not available through a standard Python library.
  \item The \texttt{utils} folder contains some utility files for simulation, experimentation and visualization.
  \item At the root folder we have files for executing our experiments, which accept many user-specified parameters (e.g. number of iterations, type of data set, \ldots). In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while script the \texttt{results\_all.py} allows the execution of our Bayesian methods without a cross-validation loop.
\end{itemize}

When possible, the code was implemented in a generic way that would allow for easy extensions or derivations. It was also developed with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy}\footnote{\url{https://numpy.org}} and \textit{scipy} libraries. Moreover, we followed closely the style of the \textit{scikit-learn} and \textit{scikit-fda} libraries, so our methods are fully compatible and could be integrated (after some minor tweaking) with both of them. The code for the experiments was executed in a machine with 4 cores, and a random seed with value 2022 was set for reproducibility. We provide a script file \texttt{launch.sh} that illustrates a typical execution.

Lastly, there are also Jupyter notebooks that demonstrate the use of our methods in a more visual way. Inside these notebooks there is a step-by-step guide on how one might execute our algorithms, accompanied by many graphical representations, and offering the possibility of changing multiple parameters to experiment with the code. In addition, there is also a notebook that can be used to generate all the tables and figures of this document pertaining to the experimental results.

### 2022-09-19 chapters/background-related-work.tex
Small changes after revision

FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}. Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint. Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}. At the same time, much effort has also been put into the task of building a sound theoretical basis for FDA, generalizing different concepts to the infinite-dimensional framework. Examples of this endeavor include the definition of centrality measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep{berrendero2020mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others. Interestingly enough, these last two are examples of situations in which a RKHS-based approach provides useful insights.

@@

In spite of the apparent generality of model~\eqref{eq:l2-linear-model}, it can be shown that it is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals of the process, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this. It turns out that in both linear and logistic scenarios a natural and more general alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space. As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions. In fact, it does include a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity. These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional} for the case of functional logistic regression.

@@

The term Bayesian inference usually refers to a wide class of methods that to some extent employ Bayes' theorem to update the initial probability assigned to a hypothesis (or a parameter) when new information is available. It can be seen as a general tool for modeling situations or problems that involve uncertainty; some examples include Bayesian hierarchical modeling, Bayesian regression or even Bayesian neural networks \citep[see e.g.][]{murphy2012machine, bishop2006pattern}. Specifically, in this work we will be interested in performing parameter estimation in a Bayesian framework, so that pre-existing information and beliefs about the parameters can be incorporated into the model.

@@

At this point, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\), which informally amounts to adding all the limits of Cauchy sequences in \(\Hcal_0(K)\), turning it into a genuine Hilbert space with the inner product extended accordingly. Then it is immediate to see that the reproducing property is retained. An important consequence is that \(\Hcal(K)\) is a space of actual functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.

Indeed, another way of defining a RKHS is via the continuity of all the \textit{evaluation operators}, i.e., \(\delta_t(f) := f(t)\) for \(f\in\Hcal(K)\) and \(t\in[0,1]\). In this case, the interpretation is that if two functions are close in the RKHS norm, they are also pointwise close. Although in this definition there is no explicit mention of the kernel, it can be recovered through the Riesz representation theorem. We summarize below some interesting properties of \(\Hcal(K)\).

@@

\begin{proposition}
  For every \(t\in[0,1]\) and \(f, g \in \Hcal(K)\), the following relations hold:
\begin{enumerate}
  \item \(\dotprod{X}{K(t, \cdot)}_K = X(t)\), a particular reproducing property.
  \item \(\E[\dotprod{X}{f}_K] = 0\).
  \item \(\E[\dotprod{X}{f}_K\dotprod{X}{g}_K] = \dotprod{f}{g}_K\).
\end{enumerate}
\end{proposition}
Note that the statement \textit{(iii)} above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.

@@

Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning. For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property. First, a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS. Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
\[
K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\Hcal}.
\]
The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with a simple expression but a possibly complex (e.g. infinite-dimensional) feature space, such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).

@@

Although this algorithm performs well on a wide range of problems, it can be difficult to find a suitable jump distribution when the number of dimensions is high. If we know the conditional distribution of each variable on the rest of them, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension. Specifically, in each step we get unidimensional samples \textit{in turn} (e.g. via rejection sampling) of each variable conditional on the rest of them, using always the more up-to-date values available:
\[
\pi(x_l^{(i+1)} \mid x_{1}^{(i+1)}, \dots, x_{l-1}^{(i+1)}, x_{l+1}^{(i)}, \dots, x_L^{(i)}), \quad l=1,\dots,L.
\]

@@

Generally speaking, a MCMC algorithm can be described through a function \(R\) as \(\Lambda(t+1)=R(\Lambda(t), \xi(t), \pi)\), where \(\Lambda(t)\) is the state of the chain at instant \(t\), \(\pi\) is the objective distribution, and \(\xi(t)\) is a sequence of i.i.d. random variables that represent the random behavior of the chain. With this notation, the  affine-invariance property can be characterized as \citep{goodman2010ensemble}:
 \[
 R(A\lambda+b, \xi(t), \pi_{A,b}) = AR(\lambda, \xi(t), \pi) + b,
 \]
 for all \(A,b\) and \(\lambda\), and for almost all \(\xi(t)\). This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Lambda(t)+b\) for all \(t\). In \citet{goodman2010ensemble} the authors consider an ensemble of samplers with the affine invariance property. Specifically, they work with a set \(\Lambda=(\Lambda_1, \dots, \Lambda_L)\) of \textit{walkers}, where \(\Lambda_l(t)\) represents an individual chain at time \(t\). At each iteration, an affine-invariant transformation is used to find the next point, which is constructed using the current values of the rest of the walkers (similar to Gibb's algorithm), namely the \textit{complementary ensemble}
 \[
  \Lambda_{-l}(t) = \{\Lambda_1(t+1), \dots, \Lambda_{l-1}(t+1), \Lambda_{l+1}(t), \dots, \Lambda_L(t)\}, \quad l=1,\dots, L.
  \]

@@

  \paragraph{Walk move.} For each walker \(1\leq l \leq L\) a random subset \(S_l \subseteq \Lambda_{-l}(t)\) with \(|S_l| \geq 2\) is chosen, and the proposed move is
\[
\Lambda_l(t) \to \Gamma = \Lambda_l(t) + W,
\]
where \(W\) is a normal distribution with mean \(0\) and the same covariance as the sample covariance of all walkers in \(S_l\). The acceptance probability in this case is just the Metropolis ratio, i.e., \(\alpha=\min\{1, \pi(\Gamma)/\pi(\Lambda_l(t))\}\).

>>>>

FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}. Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint. Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}. At the same time, much effort has also been put into the task of building a sound theoretical basis for FDA, generalizing different concepts to the infinite-dimensional framework. Examples of this endeavor include the definition of centrality measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep[e.g.][]{berrendero2020mahalanobis, galeano2015mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others. Interestingly enough, these last two are examples of situations in which a RKHS-based approach provides useful insights.

@@

In spite of the apparent generality of model~\eqref{eq:l2-linear-model}, it can be shown that it is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals of the process, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this. It turns out that in both linear and logistic scenarios a natural and more general alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space. As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions. In fact, it does include a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity. These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional} for the case of functional logistic regression.

@@

The term Bayesian inference usually refers to a wide class of methods that to some extent employ Bayes' theorem to update the initial probability assigned to a hypothesis (or a parameter) when new information is available. It can be seen as a general tool for modeling situations or problems that involve uncertainty; some examples include Bayesian hierarchical modeling, Bayesian regression or even Bayesian neural networks \citep[see e.g.][]{murphy2012machine, bishop2006pattern}. Specifically, in this work we will be interested in performing parameter estimation in a Bayesian framework, so that pre-existing information and beliefs about the parameters can be incorporated into the model in question.

@@

At this point, \(\Hcal(K)\) is defined to be the completion of \(\Hcal_0(K)\) under the norm induced by the scalar product \(\dotprod{\cdot}{\cdot}_K\), which informally amounts to adding all the limits of Cauchy sequences in \(\Hcal_0(K)\), turning it into a genuine Hilbert space with the inner product extended accordingly. Then, it is immediate to see that the reproducing property is retained. An important consequence is that \(\Hcal(K)\) is a space of actual functions and not of equivalence classes, since the values of the functions at particular points are in fact relevant, unlike in \(L^2\)-spaces.

Indeed, another way of characterizing a RKHS is via the continuity of all the \textit{evaluation operators}, i.e., \(\delta_t(f) := f(t)\) for \(f\in\Hcal(K)\) and \(t\in[0,1]\). In this case, the interpretation is that if two functions are close in the RKHS norm, they are also pointwise close. Although in this definition there is no explicit mention of the kernel, it can be recovered through the Riesz representation theorem. We summarize below some interesting properties of \(\Hcal(K)\).

@@

\begin{proposition}
  For every \(t\in[0,1]\) and \(f, g \in \Hcal(K)\), the following relations hold:
\begin{enumerate}
  \item \(\dotprod{X}{K(t, \cdot)}_K = X(t)\), a particular reproducing property.
  \item \(\E[\dotprod{X}{f}_K] = 0\).
  \item \(\E[\dotprod{X}{f}_K\dotprod{X}{g}_K] = \dotprod{f}{g}_K\).
\end{enumerate}
\end{proposition}
Note that statement \textit{(iii)} above provides yet another characterization of the inner product in \(\Hcal(K)\), where \(\dotprod{X}{f}_K\) and \(\dotprod{X}{g}_K\) are understood as the random variables representing \(f\) and \(g\) in \(\Lcal(X)\), respectively.

@@

Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning. For example, the well-known \textit{kernel trick} can be seen as a specific usage of the reproducing property. First, a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS. Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
\[
K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\Hcal}.
\]
The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with a simple expression but a possibly complex (e.g. infinite-dimensional) feature space, such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).

@@

Although this algorithm performs well on a wide range of problems, it can be difficult to find a suitable jump distribution when the number of dimensions is high. If we know the distribution of each variable conditional on the rest of them, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension. Specifically, in each step we get unidimensional samples \textit{in turn} (e.g. via rejection sampling) of each variable conditional on the rest of them, using always the more up-to-date values available:
\[
\pi(x_l^{(i+1)} \mid x_{1}^{(i+1)}, \dots, x_{l-1}^{(i+1)}, x_{l+1}^{(i)}, \dots, x_L^{(i)}), \quad l=1,\dots,L.
\]

@@

Generally speaking, a MCMC algorithm can be described through a function \(R\) as \(\Lambda(t+1)=R(\Lambda(t), \xi(t), \pi)\), where \(\Lambda(t)\) is the state of the chain at instant \(t\), \(\pi\) is the objective distribution, and \(\xi(t)\) is a sequence of i.i.d. random variables that represent the random behavior of the chain. With this notation, the  affine-invariance property can be characterized as \citep{goodman2010ensemble}:
 \[
 R(A\lambda+b, \xi(t), \pi_{A,b}) = AR(\lambda, \xi(t), \pi) + b,
 \]
 for all \(A,b\) and \(\lambda\), and almost all \(\xi(t)\). This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Lambda(t)+b\) for all \(t\). In \citet{goodman2010ensemble} the authors consider an ensemble of samplers with the affine invariance property. Specifically, they work with a set \(\Lambda=(\Lambda_1, \dots, \Lambda_L)\) of \textit{walkers}, where \(\Lambda_l(t)\) represents an individual chain at time \(t\). At each iteration, an affine-invariant transformation is used to find the next point, which is constructed using the current values of the rest of the walkers (similar to Gibb's algorithm), namely the \textit{complementary ensemble}
 \[
  \Lambda_{-l}(t) = \{\Lambda_1(t+1), \dots, \Lambda_{l-1}(t+1), \Lambda_{l+1}(t), \dots, \Lambda_L(t)\}, \quad l=1,\dots, L.
  \]

@@

  \paragraph{Walk move.} For each walker \(1\leq l \leq L\) a random subset \(S_l \subseteq \Lambda_{-l}(t)\) with \(|S_l| \geq 2\) is selected, and the proposed move is
\[
\Lambda_l(t) \to \Gamma = \Lambda_l(t) + W,
\]
where \(W\) is a normal distribution with mean \(0\) and the same covariance as the sample covariance of all walkers in \(S_l\). The acceptance probability in this case is just the Metropolis ratio, i.e., \(\alpha=\min\{1, \pi(\Gamma)/\pi(\Lambda_l(t))\}\).

### 2022-09-19 chapters/bayesian-functional-regression.tex
Small changes after revision

At this point we propose to follow a Bayesian approach to estimate the parameters of the model, which we believe is in line with the idea of simplicity we pursue, and also introduces an additional layer of flexibility in the model. In this way, we can include problem-specific information in the model through the use of prior distributions, and on top of that, this method works almost unaltered for both linear and logistic regression models. The general idea will be to impose a prior distribution on the functional parameter to eventually derive a posterior model after incorporating the available sample information.

In view of~\eqref{eq:alpha_h0}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_0(K)\)) it suffices to first consider a discrete distribution on \(p\), and then impose \(p\)-dimensional continuous prior distributions on the coefficients \(\beta_j\) and the times \(t_j\) given \(p\). Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods). Moreover, note that starting from a probability distribution \(\P_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\P\) on \(\Hcal(K)\) merely by defining \(\P(B) = \P_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\). Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).

@@

When \(p\) is fixed, the parameter space becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this \((2p+2)\)-dimensional space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\). Before proceeding further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\D_n\). For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
\begin{equation}\label{eq:rkhs-model-linear-2}
  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
\end{equation}

@@

Besides, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters. In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(l^1\)-type loss function while the former implicitly optimizes an \(l^2\) loss. On the other hand, the mode is also a good candidate because it represents the point of highest probability density. At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.

@@

In much the same way as the linear regression model described above, this RKHS-based logistic regression model offers some advantages over the \(L^2\)-model~\eqref{eq:l2-logistic-model}. First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below. Secondly, it can be shown that under mild conditions the general RKHS functional logistic model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes, and in this case it also entails the \(L^2\)-model \citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for the reduced model. Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well. Indeed, the use of a finite-dimensional approximation  mitigates the problem of non-existence of the MLE in the functional case. However, let us recall that even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}, though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.

@@

As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model. However, in this case the nuisance parameter \(\sigma^2\) only appears as part of the hierarchical prior distribution, and not in the final model. The posterior distribution is then derived after a routine calculation.

@@

With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
\begin{equation}\label{eq:summarize-predict-logistic}
\hat Y_i = \I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
\end{equation}
where \(\I\) is the indicator function (\(\I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise). The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions. On the other hand, the prediction method that takes into account the entire posterior approximation (that is, the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples. Hence, there are primarily two possible outcomes.

  \paragraph{Average sampled probability.} If we choose to average the approximate probabilities \(p_i^{(m)*} = \P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
  \[
    \hat Y_i = \I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
  \]
  \paragraph{Average sampled response.} Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
  \[
    \hat Y_i = \I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
  \]
  In this case, each \(Y_i^{(m)*}\) is a random variable that follows a Bernoulli distribution with parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\). Note that when \(\gamma=1/2\) this is equivalent to predicting \(Y_i\) according to the majority vote of all the \(Y_i^{(m)*}\).\\

Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.

>>>>

At this point we propose to follow a Bayesian approach to estimate the parameters of the model, which we believe is in line with the idea of simplicity we pursue, and also introduces an additional layer of flexibility into the model. In this way, we can include problem-specific information through the use of prior distributions, and on top of that, this method works almost unaltered for both linear and logistic regression models. The general idea will be to impose a prior distribution on the functional parameter to eventually derive a posterior distribution after incorporating the available sample information.

In view of~\eqref{eq:alpha_h0}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_0(K)\)) it suffices to first consider a discrete distribution on \(p\), and then impose \(p\)-dimensional continuous prior distributions on the coefficients \(\beta_j\) and the times \(t_j\) given \(p\). Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods). Moreover, note that starting from a probability distribution \(\P_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\P\) on \(\Hcal(K)\) merely by defining \(\P(B) = \P_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\). Consequently, our simplifying assumption on \(\alpha\) is not actually very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\).

@@

When \(p\) is fixed, the parameter space becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this \((2p+2)\)-dimensional space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\). Before proceeding further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\D_n\). Indeed, for \(\theta \in \Theta_p\) the reinterpreted model assumes the form
\begin{equation}\label{eq:rkhs-model-linear-2}
  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
\end{equation}

@@

Besides, observe that the choice of a specific point estimator to summarize the posterior distribution results in a veiled assumption of an underlying loss function between the estimated and real parameters. In general, the mean is more sensitive to outliers and the median is more robust, but the latter assumes an \(l^1\)-type loss function while the former implicitly optimizes an \(l^2\) loss. On the other hand, the mode is also a good candidate because it represents the point of highest probability density, following a \textit{maximum a posteriori} (MAP) approach. At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.

@@

In much the same way as the linear regression model described above, this RKHS-based logistic regression model offers some advantages over the \(L^2\)-model~\eqref{eq:l2-logistic-model}. First and foremost, it has a more straightforward interpretation and allows for a workable Bayesian approach, as we will demonstrate below. Secondly, it can be shown that under mild conditions the general RKHS functional logistic model holds whenever the conditional distributions \(X | Y=0\) and \(X|Y=1\) are homoscedastic Gaussian processes, and in some cases it also entails the \(L^2\)-model \citep[see Theorem 1 in][]{berrendero2018functional}; this provides a sound theoretical motivation for the reduced model. Furthermore, a maximum likelihood approach for parameter estimation (although not considered here) is possible as well. Indeed, the use of a finite-dimensional approximation  mitigates the problem of non-existence of the MLE in the functional case. However, let us recall that even in finite-dimensional settings there are cases of quasi-complete separation in which the MLE does not exist \citep{albert1984existence}, though this issue can be circumvented using, for example, Firth's corrected estimator \citep{firth1993bias}.

@@

As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model. However, in this case the nuisance parameter \(\sigma^2\) only appears as part of the hierarchical prior distribution, and not in the final model. The posterior distribution is again derived after a routine calculation.

@@

With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
\begin{equation}\label{eq:summarize-predict-logistic}
\hat Y_i = \I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
\end{equation}
where \(\I\) is the indicator function (\(\I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise). The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions. On the other hand, the prediction method that takes into account the entire posterior approximation (i.e. the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples. Hence, there are primarily two possible outcomes.

  \paragraph{Average sampled probability.} If we choose to average the approximate probabilities \(p_i^{(m)*} = \P(Y_i =1 | X_i,\theta^{(m)*})\) computed following~\eqref{eq:rkhs-model-logistic-2-parameter}, the resulting predictor is
  \[
    \hat Y_i = \I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
  \]
  \paragraph{Average sampled response.} Deciding to average the approximate binary responses \(Y_i^{(m)*}\) instead (see~\eqref{eq:sampled-response-vector}) leads to computing the predictions as
  \[
    \hat Y_i = \I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
  \]
  In this case, each \(Y_i^{(m)*}\) is a random variable that follows a Bernoulli distribution with parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\). Note that when \(\gamma=1/2\) this is equivalent to predicting \(\hat Y_i\) according to the majority vote of all the \(Y_i^{(m)*}\).\\

Lastly, the variable selection method is essentially the same as in the case of functional linear regression: we select \(p\) time instants from each trajectory based on a summary of the posterior distribution \(\tau | \D_n\), and then feed the reduced data set to a finite-dimensional binary classification procedure.

### 2022-09-19 chapters/conclusions.tex
Small changes after revision

\begin{itemize}
  \item To delve further into the relationship between RKHS's and functional data problems, a connection that has proven to be fruitful in many scenarios. A first idea would be to extend the RKHS-based logistic regression model to a generalized functional linear model with an arbitrary link function.
  \item To try to derive some theoretical properties of our Bayesian predictors. For example, consistency and/or robustness results regarding the posterior distribution would be an excellent complement to the practical side of this work.
  \item To find other prior distributions for our parameters that perform better in general, or to eliminate the need of the hyperparameters \(b_0\) or \(\eta\).
  \item To experiment with other MCMC algorithms for posterior approximation. Specifically, it would be very interesting to implement an efficient and reliable reversible-jump MCMC method in Python, which as we have already mentioned is a better fit for our particular Bayesian model.
\end{itemize}

>>>>

\begin{itemize}
  \item To delve further into the relationship between RKHS's and functional data problems, a connection that has proven to be fruitful in many scenarios. A first idea would be to extend the RKHS-based logistic regression model to a generalized functional linear model with an arbitrary link function.
  \item To try to derive some theoretical properties of our Bayesian predictors. For example, consistency and/or robustness results regarding the posterior distribution would be an excellent complement to the practical side of this work.
  \item To find other prior distributions for our parameters that perform better in general, or to eliminate the need of the hyperparameters \(b_0\) or \(\eta\).
  \item To experiment with other MCMC algorithms for posterior approximation. Specifically, it would be interesting to implement an efficient and reliable reversible-jump MCMC method in Python, which as we have already mentioned is a better fit for our particular Bayesian model.
\end{itemize}
